JD notes on effects

### 2022-06-18 manuscript/body.tex
Gone through all BB's comments?

Both predictions and effects tell us what we expect the response to be for a given value of \emph{focal} input variable (scientific variable of interest). The major difference between the two lies in the construction of the associated confidence intervals. In particular, for predictions, we want to use the classic curved confidence intervals (or prediction intervals) which capture the uncertainty in the whole model and generally focus on the \emph{total} effects, and so will often be suited for a univariate model. On the other hand, effects, attempt to show the effect of a focal input variable. In this case, we suggest the narrower confidence intervals which focus on uncertainty in the effect of the focal input variable only, and so will often be suited for the direct effects – after controlling for other non-focal input variables – from a multivariate model.

@@

The most common way of choosing the representative values is taking unique levels of the focal input variable if discrete or taking appropriately sized quantiles if continuous. The predictions are then calculated by holding the non-focal input variables at their reference point (values chosen for the non-focal input variables, e.g., means), while varying the focal input variable, with the goal that the estimates represent how the model responds to the changes in the focal input variable \citep{fox2009effect, hanmer2013behind}. This procedure generates -- \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to these quantities as predictions and effects since they should, for example, tell us what we would expect the response to be at a particular value or level of the focal input variable, for an ``average case''. . As mentioned above, in our case, the key difference between predictions and effects is in the computation of the confidence intervals.

@@

\caption{{\bf A comparison of \pkg{emmeans}, \pkg{effects} and \pkg{varpred} predictions for $x_1$ on $y$ for models with and without interactions.}
The horizontal blue, green and black lines are the mean predictions, i.e., $\bar{\hat{y}}$, the red ones are the means of the simulated $y$, i.e., $\bar{y}$, while the vertical dotted grey lines are the means of the corresponding focal input variable. The grey points are the binned simulated $y$. The trend lines represent the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other input variables at their average for the three methods and the true predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated value in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches closely match the true predictions. B: Even with the simple interaction between non-focal input variables, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and true predictions, respectively; in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not the proposed \pkg{varpred}.}

@@

In the absence of interaction (Equation~\ref{eq:simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, Fig~\ref{fig:justify_plots}A. However, in the presence of interaction, even as simple as the one in Equation~\ref{eq:simple_inter_higher}, the estimates starts to differ. In particular, \pkg{emmeans} and \pkg{effects} give similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), Fig~\ref{fig:justify_plots}B. To generate Fig~\ref{fig:justify_plots}A,  \pkg{emmeans} and \pkg{effects} average $x_2$ and $x_3$, i.e., the input variables while \pkg{varpred} averages the predictor variables corresponds to $x_2$ and $x_3$. On other hand, to generate Fig~\ref{fig:justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction term ($x_2x_3$). In particular \pkg{emmeans} and \pkg{effects} use input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses predictor variables to compute $\over{x_2x_3}$. In the case of Fig~\ref{fig:justify_plots}A, averaging the input variables is equivalent to averaging the predictor variables.

@@

\begin{figure}
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predictions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal input variable, i.e., the model center, corresponds to the effects of the focal input variable from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal input variables. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal input variable.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence bands are much wider because they include uncertainties associated with the intercept and non-focal input variables, but narrower and crosses at the mean of the focal input variable in \pkg{varpred}. We refer to this point as the \emph{center point} and the approach as \emph{mean-based}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal input variable we are more certain about, i.e., mean of the predictor variable corresponding to the focal input variable. For simple models, the point where the confidence bands crosses is the model center; it corresponds to anchor we choose, in this case, the mean of the focal input variable, but we generally choose the mean of the focal predictor (predictors corresponding to the focal input variable).

@@

Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor. The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only and depend on the anchor.

The main purpose of this article is to discuss and implement various approaches for computing predictions and provide an alternative method for computing the associated confidence intervals, i.e., effects. We further explore and demonstrate, using simulated data, approaches for correcting bias in predictions for GL(M)Ms involving nonlinear link functions.

@@

\item \textbf{Focal and non-focal input variables}: Focal input variable refers to the input variable we are interested in. For multivariate models, all other remaining input variables are non-focal input variables. For example, consider model described by Equation~\ref{eq:simple_inter_higher_no_interaction}. Suppose we are interested in the effect of $x_1$ on the predicted $y$, then $x_1$ is the focal input variable while $x_2$ and $x_3$ are non-focal non-focal variables.

@@

\item \textbf{Non-focal predictors:} All columns of the model matrix associated with the non-focal input variable(s). For example, in the previous household size example, all the columns associated with religion are non-focal predictors.
\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more predictor variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of predictor variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots} (with interaction). In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for center point. If we have more than one predictor variables associated with the focal input variable, however, the model center does not correspond to any given value of the variable, and we don't get the nice crossing we get in simpler cases. See Fig~\ref{fig:pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal input variables, when estimating the predictions. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate CIs becomes important. This is not true for the more conventional prediction CIs. We can choose any anchor, e.g., zero-anchored (the value of the anchor is $0$), but center point seems like a natural, stable choice, we call this center-anchored. In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the minimum and center (mean) of the focal input variable.

@@

\begin{figure}
\begin{center}
\includegraphics{justify_anchors.combined.pdf}
\end{center}
\caption{{\bf A comparison of min-based and center- (mean-) based anchors.} The center-based effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

Consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function (in the case of simple linear model), where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the prediction with respect to the focal predictor in question \citep{fox2009effect}.

@@

Higher order terms such as interactions, splines, polynomials, etc., can be within the focal input variable, within (between) non-focal input variable(s) or between the focal and non-focal input variable(s). To distinguish the three, suppose the model which describes the hypothetical simulated household size simulation based on a number of socio-demographic factors such as age and wealthindex is

\begin{align}\label{eq:lm_cubic}
\mathrm{hh~size}_i &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i \nonumber\\
\end{align}

In the first case, with \code{Age} as the focal input variable, with cubic polynomial interaction ($3$ focal predictors associated with it). In this case, each of the focal predictors ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are evaluated independently across the chosen levels of the focal input variable, $Age_i$. Specifically, the higher order terms are treated as additional columns of the model matrix evaluated with the same values of the focal input variable, while non-focal predictors are fixed at their reference point as discussed in the previous section. In this case, the predictions associated with \code{Age} on the linear predictor scale become

\begin{align*}
\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i\\
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}}
\end{align*}

In the second case, with \code{Wealthindex} as the focal input variable, the non-focal variable, \code{Age}, is a cubic polynomial. In this case, the non-focal predictors ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are simply treated as additional columns in the predictor space and an appropriate choice of reference point applies just like in the models without higher non-focal variable interactions. For instance, in mean-based approach, we average all non-focal predictors. Thus

\begin{align*}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age\mathop{\vphantom{^2}}}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age\mathop{\vphantom{^2}}}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i
\end{align*}

Lastly, consider model described by Equation~\ref{eq:simple_inter_higher}, previously used to demonstrate how to handle interactions in non-focal input variables. Here, we use the same model to demonstrate how to handle interactions between the focal input variable, $x_2$, and non-focal input variable, $x_3$. In this case, the non-interacting predictors are treated as non-focal predictors, with appropriate reference points. However, for the interacting predictors, we first choose representative values for the focal input variable and some particular values of the interacting non-focal input variable in a similar way we choose values for the focal input variable or predetermined. In our example, suppose we pick $i$ and $j$ unique values of the focal input variable, $x_2$ and interacting input variable $x_3$, respectively. The prediction on linear predictor scale is given by

\begin{align*}
\eta^\star(x_{2i}, x_{3j}, {\bar{x}^\star_1}) = \beta_0 + \beta_1 \bar{x}^\star_1 + \beta_2x_{2i} + \beta_3x_{3j} + \beta_{23}x_{2i}x_{3j}.
\end{align*}

@@

Let $\bxo$ be a centered model matrix previously defined, and let $\ba$ be an anchor matrix, with the same dimensions and entries in all non-focal predictors as $\bxo$. Let $\baf$ be the column of $\ba$ corresponding to focal predictor(s) defined in $\bxo$. Any appropriate values can be chosen for $\baf$ but for model center (center-anchored), we use $\bafc$ which is the mean of the focal predictor(s). Thus 
%
\begin{align}\label{eq:centered_variance}
\boldsymbol\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top).
\end{align}
%
We can see that $\forall ~\bxof=\baf$, $\bxo - \baf = \boldsymbol 0$, hence $\boldsymbol\sigma_f^2 = \boldsymbol{0}$. Similarly, for all values of $\bxof$ close to anchor point $\bafc$, the term $(\bxo - \bafc)$ and $\boldsymbol\sigma_f^2$ goes to $\boldsymbol 0$ and $\boldsymbol\sigma_f^2 = \boldsymbol 0$ if $\bxof=\bafc$. This means that $\boldsymbol\sigma_f^2$ close to the anchor point are smaller than those away from the anchor point; and results to confidence intervals which are narrower around the anchor or crosses at the anchor point for simple models. This is the effect of the particular focal input variable. By setting $\baf = \boldsymbol 0$, we get the variances for the conventional CIs, Equation~\ref{eq:conventional_variance}.

@@

An alternative approach to choosing a reference point is to make predictions over all observations of the non-focal input variables (members of the population) \citep{hanmer2013behind}. The nonlinear transformation involved in these computations is always \emph{one-dimensional}; all of the multivariate computations required are at the stage of collapsing the multidimensional set of predictors for some subset of the population to a one-dimensional distribution of $\eta^\star(x_f, \nset{x})$, which is a function of the chosen values of the focal predictor and the whole sample of non-focal predictors, as opposed to the definition in Equation~\ref{eq:eta_mean}. More specifically:
\begin{itemize}
\item compute linear predictor associated with whole sample of the non-focal predictors, $\nset{\eta} = \sum \nset{\beta} \nset{x}$
\item compute linear predictor associated with the chosen values focal input variable, $\eta_{jf} = \sum{\beta_f x_{jf}}$
\item for every value of the focal linear predictor, $\eta_{jf}$, compute

\begin{align}\label{eq:pop_eta} 
\eta_j^\star(\eta_{jf}, \nset{\eta})  &= \eta_{jf} + \nset{\eta} \nonumber \\
&= \eta_j^\star(x_f, \nset{x})
\end{align}
\end{itemize}

Once Equation~\ref{eq:pop_eta} is computed, we back-transform the estimates to the original scale and average over the levels of the focal predictors, $j$:

\begin{align}\label{eq:pop_response} 
\hat{y}_f  &= \textrm{mean} ~ g^{-1} \left(\eta_j^\star(x_f, \nset{x})\right)
\end{align}

We make similar adjustments to compute the variances of the predictions at every level of the focal input variable:

\begin{align}
\sigma_{jf}^2 = \textrm{Diag}(\bX^\star_{jc} \boldsymbol{\Sigma} \bX^{\star\top}_{jc})
\end{align}
where $\bX^{\star}_{jc} = \{\bX_{jf}^\star, \nset{{\bX}^\star} - \nset{{\bar{\bX}}^\star}\}$ and

\begin{align}
\mathrm{CI}_f = \mathrm{mean} ~ g^{-1} \left(\eta_j^\star(x_f, \nset{x}) \pm q\sigma_{jf}\right)
\end{align}

For models with random effects components, we make further adjustment to correct for bias induced by the random effects terms. In whole-sample approach, we treat the random effects terms as additional non-focal predictors and simply make adjustment to Equation~\ref{eq:pop_eta}. In particular

\begin{align}\label{eq:pop_eta_re} 
\tau &= \bZ b \nonumber \\
\eta_j^\star(x_f, \nset{x}, \tau)  &= \eta_j^\star(x_f, \nset{x}) + \tau
\end{align}
where $\bZ$ and $b$ are the design matrix and a vector of random effects, respectively.

@@

To illustrate the distinction between the two approaches, we simulate a multivariate model described in \nameref{S1_Appendix}. We also use the same model to illustrate how predictions and effects differ in the presence of higher order interactions in focal or non-focal input variables. We then generated two sets of mean-based predictions and effects: 1) one with \code{age} as the focal input variable (a cubic polynomial focal predictor); and 2) one with \code{Wealthindex} as the focal in put variable (linear focal input variable). In both cases the, the center point was used as the anchor for the effects. Then compared the predictions and effects as shown in Fig~\ref{fig:pred_cubic_plots}.

>>>>

Both predictions and effects tell us what we expect the response to be for a given value of \emph{focal} input variable (scientific variable of interest). The major difference between the two lies in the construction of the associated confidence intervals. In particular, for predictions, we want to use the conventional curved confidence intervals (or prediction intervals) which capture the uncertainty in the whole model and generally focus on the \emph{total} effects, and so will often be suited for a univariate model. On the other hand, effects, attempt to show the effect of a focal input variable. In this case, we suggest the narrower confidence intervals which focus on uncertainty in the effect of the focal input variable only, and so will often be suited for the direct effects – after controlling for other non-focal input variables – from a multivariate model.

@@

The most common way of choosing the representative values is taking unique levels of the focal input variable if discrete or taking appropriately sized quantiles if continuous. The predictions are then calculated by holding the non-focal input variables at their reference point (values chosen for the non-focal input variables, e.g., means), while varying the focal input variable, with the goal that the estimates represent how the model responds to the changes in the focal input variable \citep{fox2009effect, hanmer2013behind}. This procedure generates -- \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to these quantities as predictions and effects since they should, for example, tell us what we would expect the response to be at a particular value or level of the focal input variable, for an ``average case''. . As mentioned above, in our case, the key difference between predictions and effects is in the computation of the confidence intervals. The concept of effects has been presented in the literature -- Shi et al. \citep{shi_evidence_2017} used effect-styled plots to compare the difference in sexual risk behaviors between circumcised and uncircumcised men.

@@

\caption{{\bf A comparison of \pkg{emmeans}, \pkg{effects} and \pkg{varpred} predictions for regression of $x_1$ on $y$.} The horizontal dashed lines are the mean of the predicted and simulated $y$, $\bar{\hat{y}}$ and $\bar{y}$, respectively. The grey points are the binned simulated $y$. The trend lines represent the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other input variables at their mean and the true predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated value in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches match the true predictions. B: Even with the simple interaction between non-focal input variables, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and true predictions, respectively; in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not the proposed \pkg{varpred}.}

@@

In the absence of interactions (Equation~\ref{eq:simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, Fig~\ref{fig:justify_plots}A. However, in the presence of interactions, even as simple as the one in Equation~\ref{eq:simple_inter_higher}, the estimates start to differ. In particular, \pkg{emmeans} and \pkg{effects} give similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), Fig~\ref{fig:justify_plots}B. To generate Fig~\ref{fig:justify_plots}A,  \pkg{emmeans} and \pkg{effects} average $x_2$ and $x_3$, i.e., the input variables while \pkg{varpred} averages the predictor variables corresponds to $x_2$ and $x_3$. On other hand, to generate Fig~\ref{fig:justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction term ($x_2x_3$). In particular \pkg{emmeans} and \pkg{effects} use input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses predictor variables to compute $\over{x_2x_3}$. In the case of Fig~\ref{fig:justify_plots}A, averaging the input variables is equivalent to averaging the predictor variables.

@@

\begin{figure}
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf The prediction-styled and effect-styled $95\%$ confidence intervals.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence intervals around the predictions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal input variable, i.e., the center point, corresponds to the effects of the focal input variable from \pkg{varpred}. For simple models, the effect-styled confidence bands crosses at the center point (an anchor). The prediction-styled curves incorporates not only the uncertainties due to intercept term but also other non-focal input variables. On the other hand, effect-styled curves only take into account the main effect uncertainty of the focal input variable.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence intervals are much wider because they include uncertainties associated with the intercept and non-focal input variables, but narrower and crosses at the mean of the focal input variable in \pkg{varpred}. We refer to this point as the \emph{center point} and the approach as \emph{mean-based}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal input variable we are more certain about, i.e., the center point. For simple models, the point where the confidence intervals curves cross is the model center or simply center point; it corresponds to anchor we choose, in this case, the center point (which is generally an appropriate choice). Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor. The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only and depend on the anchor.

@@

The main purpose of this article is to discuss and implement various approaches for computing predictions and provide an alternative method for computing the associated confidence intervals, i.e., effects. We further explore and demonstrate, using simulated data, approaches for correcting bias in predictions for GL(M)Ms involving nonlinear link functions. The proposed method and \proglang{R} software package will complement the existing ones by providing: 1) a straightforward way to generate effects, and 2) an alternative and a more robust way to correct for predictions bias in GL(M)Ms.

@@

\item \textbf{Focal and non-focal input variables}: Focal input variable refers to the input variable we are interested in. For multivariate models, all other variables other than focal input variable are non-focal input variables. For example, consider model described by Equation~\ref{eq:simple_inter_higher_no_interaction}. Suppose we are interested in the effect of $x_1$ on the predicted $y$, then $x_1$ is the focal input variable while $x_2$ and $x_3$ are non-focal input variables.

@@

\item \textbf{Non-focal predictors:} All columns of the model matrix associated with the non-focal input variable(s). For example, in the previous household size example, all the columns associated with religion are non-focal predictors if income is the focal input variable.
\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more predictor variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of predictor variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots}B. In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for the center point. If we have more than one predictor variables associated with the focal input variable, however, the model center does not correspond to any given value of the variable, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see Fig~\ref{fig:pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal input variables, when estimating the predictions. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions (for bias correction) especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the more conventional prediction confidence intervals. We can choose any anchor, e.g., zero-anchored (the value of the anchor is $0$), but center point seems like a natural, stable choice, we call this center-anchored. In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the minimum and center (mean) of the focal input variable.

@@

\begin{figure}
\begin{center}
\includegraphics{justify_anchors.combined.pdf}
\end{center}
\caption{{\bf A comparison of min- and center- anchored effects.} The center-anchored effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

Consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function, where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the prediction with respect to the focal input variable in question \citep{fox2009effect}.

@@

Higher order terms such as interactions, splines, polynomials, etc., can be within the focal input variable, within (between) non-focal input variable(s) or between the focal and non-focal input variable(s). To distinguish the three, suppose the model which describes the hypothetical simulated household size simulation based on a number of socio-demographic factors such as age and wealth index is
%
\begin{align}\label{eq:lm_cubic}
\mathrm{hh~size}_i &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i\nonumber \\
&+ \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i.
\end{align}
%
In the first case, with \code{Age} as the focal input variable, with cubic polynomial interaction ($3$ focal predictors associated with it). In this case, each of the focal predictors ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are evaluated independently across the chosen levels of the focal input variable, $Age_i$. Specifically, the higher order terms are treated as additional columns of the model matrix evaluated with the same values chosen for the focal input variable, while non-focal predictors are fixed at their reference point as discussed in the previous section. In this case, the predictions associated with \code{Age} on the linear predictor scale become
%
\begin{align}
\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i \nonumber\\
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}}.
\end{align}
%
In the second case, with \code{Wealthindex} as the focal input variable, the non-focal variable, \code{Age}, is a cubic polynomial. In this case, the non-focal predictors ($\mathrm{Age}_i, \mathrm{Age}^2_i$ and $\mathrm{Age}^3_i$) are simply treated as additional columns in the predictor space and an appropriate choice of reference point applies just like in the models without higher non-focal variable interactions. For instance, in mean-based approach, we average all non-focal predictors. Thus
%
\begin{align}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age\mathop{\vphantom{^2}}}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age\mathop{\vphantom{^2}}}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\nonumber\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i.
\end{align}
%
Lastly, consider model described by Equation~\ref{eq:simple_inter_higher}, previously used to demonstrate how to handle interactions in non-focal input variables. Here, we use the same model to demonstrate how to handle interactions between the focal input variable, $x_2$, and non-focal input variable, $x_3$. In this case, the non-interacting predictors are treated as non-focal predictors, with appropriate reference points. However, for the interacting predictors, we first choose representative values for the focal input variable and then some particular values of the interacting non-focal input variable in a similar way we choose values for the focal input variable or predetermined. In our example, suppose we pick $i$ and $j$ unique values of the focal input variable, $x_2$ and interacting input variable $x_3$, respectively. The prediction on linear predictor scale is given by
%
\begin{align*}
\eta^\star(x_{2i}, x_{3j}, {\bar{x}^\star_1}) = \beta_0 + \beta_1 \bar{x}^\star_1 + \beta_2x_{2i} + \beta_3x_{3j} + \beta_{23}x_{2i}x_{3j}.
\end{align*}
%

@@

Let $\bxo$ be a centered model matrix previously defined, and let $\ba$ be an anchor matrix, with the same dimensions and entries in all non-focal predictors as $\bxo$. Let $\baf$ be the column of $\ba$ corresponding to focal predictor(s) defined in $\bxo$. Any appropriate values can be chosen for $\baf$ but for model center (center-anchored), we use $\bafc$ which is the mean of the focal predictor(s). Thus 
%
\begin{align}\label{eq:centered_variance}
\boldsymbol\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top).
\end{align}
%
We can see that $\forall ~\bxof=\baf$, $\bxo - \baf = \boldsymbol 0$, hence $\boldsymbol\sigma_f^2 = \boldsymbol{0}$. Similarly, for all values of $\bxof$ close to anchor point $\bafc$, the term $(\bxo - \bafc)$ and $\boldsymbol\sigma_f^2$ goes to $\boldsymbol 0$ and $\boldsymbol\sigma_f^2 = \boldsymbol 0$ if $\bxof=\bafc$. This means that $\boldsymbol\sigma_f^2$ close to the anchor point are smaller than those away from the anchor point; and results to confidence intervals which are narrower around the anchor or crosses at the anchor point for simple models. This is the effect of the particular focal input variable. By setting $\baf = \boldsymbol 0$, we get the variances for the conventional CIs in Equation~\ref{eq:conventional_variance}.

@@

An alternative approach to choosing a reference point is to make predictions over all observations of the non-focal input variables (members of the population) \citep{hanmer2013behind}. The nonlinear transformation involved in these computations is always \emph{one-dimensional}; all of the multivariate computations required are at the stage of collapsing the multidimensional set of predictors for some subset of the population to a one-dimensional distribution of $\eta^\star(x_f, \nset{x})$, which is a function of the chosen values of the focal predictor and the whole sample of non-focal predictors, as opposed to the definition in Equation~\ref{eq:eta_mean}. More specifically:
\begin{itemize}
\item compute linear predictor associated with whole sample of the non-focal predictors, $\nset{\eta} = \sum \nset{\beta} \nset{x}$
\item compute linear predictor associated with the chosen values focal input variable, $\eta_{jf} = \sum{\beta_f x_{jf}}$
\item for every value of the focal linear predictor, $\eta_{jf}$, compute
%
\begin{align}\label{eq:pop_eta} 
\eta_j^\star(\eta_{jf}, \nset{\eta})  &= \eta_{jf} + \nset{\eta} \nonumber \\
&= \eta_j^\star(x_f, \nset{x}).
\end{align}
\end{itemize}
%

Once Equation~\ref{eq:pop_eta} is computed, we back-transform the estimates to the original scale and average over the levels of the focal predictors, $j$:
%
\begin{align}\label{eq:pop_response} 
\hat{y}_f  &= \textrm{mean} ~ g^{-1} \left(\eta_j^\star(x_f, \nset{x})\right).
\end{align}
%

We make similar adjustments to compute the variances of the predictions at every level of the focal input variable:
%
\begin{align}
\sigma_{jf}^2 = \textrm{Diag}(\bX^\star_{jc} \boldsymbol{\Sigma} \bX^{\star\top}_{jc})
\end{align}
where $\bX^{\star}_{jc} = \{\bX_{jf}^\star, \nset{{\bX}^\star} - \nset{{\bar{\bX}}^\star}\}$ and 
%
\begin{align}
\mathrm{CI}_f = \mathrm{mean} ~ g^{-1} \left(\eta_j^\star(x_f, \nset{x}) \pm q\sigma_{jf}\right).
\end{align}
%

For models with random effects components, we make further adjustment to correct for bias induced by the random effects terms. In whole-sample approach, we treat the random effects terms as additional non-focal predictors and simply make adjustment to Equation~\ref{eq:pop_eta}. In particular
%
\begin{align}\label{eq:pop_eta_re} 
\tau &= \bZ b \nonumber \\
\eta_j^\star(x_f, \nset{x}, \tau)  &= \eta_j^\star(x_f, \nset{x}) + \tau
\end{align}
where $\bZ$ and $b$ are the design matrix and a vector of random effects, respectively.

@@

To illustrate the distinction between the two approaches, we simulate a multivariate model described in \nameref{S1_Appendix}. We also use the same model to illustrate how predictions and effects differ in the presence of higher order interactions in focal or non-focal input variables. We then generated two sets of mean-based predictions and effects: 1) one with \code{age} as the focal input variable (a cubic polynomial focal input variable); and 2) one with \code{Wealthindex} as the focal in put variable (linear focal input variable). In both cases the, the center point was used as the anchor for the effects. Then compared the predictions and effects as shown in Fig~\ref{fig:pred_cubic_plots}.

\begin{figure}
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction and effect plots for cubic polynomial and simple focal predictors.} A: The focal input variable is a cubic polynomial. B: The focal input variable is linear, with polynomial non-focal input variable. The central curve and trend lines are the same for predictions and effects plots in each case. However, the confidence intervals for the effects are narrower than those of the predictions in both cases. The default anchor for generating effects is the center point. For simple models (or simple focal input variables), the center point is a point within the range of focal values; and therefore the confidence interval lines intersect at that that point -- B. In the case of higher order interactions such as cubic polynomials, the center point is not a point but a combination of all the predictor variables associated with the cubic polynomial, consequently, curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively.}
\label{fig:pred_cubic_plots}
\end{figure}

Predictions are the standard curves (wider); they tell us what we expect to see if we know the value of the focal input variable. In multivariate models, at times, we may want to focus on the effect of the focal input variable only. Unlike prediction plot, which captures the uncertainty in the entire model, effect plot focuses on the uncertainty of the focal input variable only. In this case, effects styled plots which are usually narrower around the anchor are preferable over the conventional prediction plots which generally focus on the \emph{total} effects.
